---
title: "Assignment ARIMA"
output: html_document
date: "2023-11-22"
---

```{r }

library(fpp)
```
```{r}
library(fpp2)
```
```{r}
library(readxl)
Gold_Price_Monthly_Dataset <- read_excel("C:/Users/mumba/Desktop/Gold_Price_Monthly_Dataset.xls")
View(Gold_Price_Monthly_Dataset)

plot(Gold_Price_Monthly_Dataset$Price, type = "l")
```
This is a line graph that is showing the price of Gold on January, 2011 till Juy, 2020


```{r}
head(x=Gold_Price_Monthly_Dataset$Date, y=Gold_Price_Monthly_Dataset$Price,type="l")
```
```{r}
#Convert into time series
ts_data_new <- ts(Gold_Price_Monthly_Dataset$Price, start = c(2011, 1), frequency = 12)
```

tdisplay shows the ACF, PACF and time series plot

```{r}
tsdisplay(ts_data_new)
```
```{r}
tsdiff <- diff(ts_data_new, differences = 1)
plot(tsdiff)
```

The diff function is employed to calculate the disparities between successive elements in a numeric vector or matrix.

```{r}
tsdisplay(tsdiff)
```
```{r}
auto_fit <- auto.arima(ts_data_new, trace=TRUE, stepwise = FALSE)
```
1. Output Summary:
The ARIMA(0,1,4) model is used for predicting future values in a time series. It doesn't rely on past values directly, but it does use the difference between consecutive values to make the data more predictable (achieve stationarity). Additionally, it considers the last four forecast errors to better capture and adjust for random fluctuations or residual errors in the data.

```{r}
auto_fit
```
auto_fit is commonly linked with functions or processes designed to automatically identify the most suitable model for a given time series. The sigma square is 2159 for the model.

```{r}
#Each attributes
attributes(auto_fit)

```
3. Perform and Plot the forecast for the next five periods.

```{r}
plot(forecast(auto_fit,h=5,level=c(99.5)))
```

The forecast for the next five periods seems to be constant with slightly decrease in the price.

2.RESIDUAL ANALYSIS

```{r}
Acf(ts_data_new)
```
```{r}
Acf(tsdiff)
```
The ACF of residual analysis for my data shows no seasonal trend. It shows the random data

```{r}
Acf(auto_fit$residuals)
```
```{r}
Box.test(residuals(auto_fit), lag=20, type="Ljung")
```
Box.test is a statistical test that helps us see if there's a pattern of repeating values in a time series. It gives us information like the chi-squared statistic, degrees of freedom, and p-value.

```{r}
plot.ts(residuals(auto_fit))
```
```{r}
hist(auto_fit$residuals)
```
```{r}
tsdiag(auto_fit)
```
The graph of standardized residuals indicates that most of them fall within the expected range, suggesting nothing unusual. However, a few outliers suggest that the ARIMA model might not fully capture all the patterns in the data from 2011 to 2020. The p-values for the Ljung-Box statistic are all higher than 1.0, indicating that the residuals don't show a pattern over time. In the ACF graph, the correlation coefficients at all time lags are within the expected range, meaning they are not statistically significant.


There is no seasonal data.

4. Show the accuracy of your ARIMA model
```{r}
accuracy(auto_fit)
```


